Impact of IPTp Regimen on Pregnancy Outcomes in a Malaria-Endemic Setting

### 2025-04-28 products/manuscript/Manuscript.qmd
Project Update

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
library(here)
here::i_am("products/manuscript/Manuscript.qmd")

# ← THIS is the one that actually sets the working dir for *all* chunks:
library(knitr)
knitr::opts_chunk$set(root.dir = here::here("products/manuscript"))

@@

**Figure 1: Distribution of Total Malaria Episodes by IPTp Treatment Arm**
```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(promo_data_clean, aes(x = total_malaria_episodes, fill = study_arm)) +
  geom_histogram(binwidth = 1, alpha = 0.6, position = "dodge") +
  labs(title = "",
       x = "Total Malaria Episodes",
       y = "Frequency",
       fill = "Treatment Arm") +
  theme_minimal()
```

@@

**Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
```{r,  echo=FALSE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(dplyr)
library(gtsummary)
library(here)

# Read in the cleaned data (if not already read)
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

# Recode variables as needed for outcomes analysis:
promo_data_clean <- promo_data_clean %>%
  mutate(
    # Create a composite adverse outcome variable: 1 if any of preterm birth, stillbirth, or low birth weight (<2.5 kg) occurs
    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0),

    # Recode Preterm Births Count: combine 0 and 1 as "1", and 2 as "2"
    PretermBirths_cat = case_when(
      preterm_births_count %in% c(0, 1) ~ "1",
      preterm_births_count == 2 ~ "2"
    ),
    PretermBirths_cat = factor(PretermBirths_cat, levels = c("1", "2"))
  )

# Create Table Z: Outcome Variables by IPTp Treatment Arm with p-values
table_outcomes <- promo_data_clean %>%
  dplyr::select(
    study_arm,
    `Malaria Infection Rate During Pregnancy` = malaria_infection_rate_during_pregnancy,
    `Placental Malaria (Rogerson Criteria)` = placental_malaria_by_rogerson_criteria,
    `Preterm Births Count` = PretermBirths_cat,
    `Stillbirth bin` = stillbirth_bin,
    `Birthweight` = birth_weight,
    `Composite Adverse Outcome` = adverse_birth_outcome
  ) %>%
  tbl_summary(
    by = study_arm,
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%
  modify_header(label = "") %>%
  modify_caption("")

# Print Table Z
table_outcomes
```

@@

**Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create the composite adverse outcome variable
promo_data_clean <- promo_data_clean %>%
  mutate(
    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0)
  )

# Fit the interaction model (ensuring that adverse_birth_outcome now exists)
model_interaction <- glm(adverse_birth_outcome ~ total_malaria_episodes * study_arm + 
                           age_at_enrollment_years + gravidity + education_level,
                         family = binomial(link = "logit"),
                         data = promo_data_clean)

# Generate predicted probabilities over the range of total malaria episodes by study arm
pred <- ggeffect(model_interaction, terms = c("total_malaria_episodes [all]", "study_arm"))

# Plot the predicted probabilities
interaction_plot <- ggplot(pred, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.2, color = NA) +
  labs(title = "",
       x = "Total Malaria Episodes",
       y = "Predicted Probability of Adverse Outcome",
       color = "Treatment Arm",
       fill = "Treatment Arm") +
  theme_minimal()

interaction_plot
```

@@

**Table 3: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
```{r,  echo=FALSE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(dplyr)
library(gtsummary)
library(here)

# Read in the cleaned data (if not already read)
promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))

# Recode variables as needed for outcomes analysis:
promo_data_clean <- promo_data_clean %>%
  mutate(
    # Create a composite adverse outcome variable: 1 if any of preterm birth, stillbirth, or low birth weight (<2.5 kg) occurs
    low_birth_weight = ifelse(birth_weight < 2.5, 1, 0),
    adverse_birth_outcome = ifelse(preterm_births_count > 0 | stillbirth_bin == 1 | low_birth_weight == 1, 1, 0),

    # Recode Preterm Births Count: combine 0 and 1 as "1", and 2 as "2"
    PretermBirths_cat = case_when(
      preterm_births_count %in% c(0, 1) ~ "1",
      preterm_births_count == 2 ~ "2"
    ),
    PretermBirths_cat = factor(PretermBirths_cat, levels = c("1", "2"))
  )

# Create Table Z: Outcome Variables by IPTp Treatment Arm with p-values
table_outcomes <- promo_data_clean %>%
  dplyr::select(
    study_arm,
    `Malaria Infection Rate During Pregnancy` = malaria_infection_rate_during_pregnancy,
    `Placental Malaria (Rogerson Criteria)` = placental_malaria_by_rogerson_criteria,
    `Preterm Births Count` = PretermBirths_cat,
    `Stillbirth bin` = stillbirth_bin,
    `Birthweight` = birth_weight,
    `Composite Adverse Outcome` = adverse_birth_outcome
  ) %>%
  tbl_summary(
    by = study_arm,
    missing = "no",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_p() %>%
  modify_header(label = "") %>%
  modify_caption("")

# Print Table Z
table_outcomes
```

@@

**Table 4: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**

@@

**Table 5: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("Images/AdjustedRatio.png")

@@

**Table 6:  Discriminative Performance (AUC) of Machine Learning Models Predicting Adverse Birth Outcomes Under Different IPTp Regimens**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("Images/Table6.Discriminative Performance (AUC) of ML Models.png")

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
# 1) Tell here where this file lives:
library(here)
here::i_am("products/manuscript/Manuscript.qmd")

# 2) Set your global chunk options (no setwd or root.dir needed):
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE
)

@@

**Figure 1: Distribution of Total Malaria Episodes by IPTp Treatment Arm**
```{r,echo=FALSE, message=FALSE, warning=FALSE}}
knitr::include_graphics("Images/Figure1.DistribTotal Malaria.png")

@@

**Table 2: Outcome Measures and Malaria Exposure Variables Stratified by IPTp Regimen**
```{r}
knitr::include_graphics("Images/Table2.Outcome Measures Malaria .png")

@@

**Figure 2: Differential Impact of IPTp Treatment on the Relationship Between Malaria Episodes and Adverse Birth Outcomes**
```{r}
knitr::include_graphics("Images/Figure2.Differential Impact of IPTp.png")
```

@@

@@

**Table 3: Interaction Between Malaria Exposure and IPTp Treatment Arm in Predicting Adverse Birth Outcomes**

@@

**Table 4: Adjusted Odds Ratios for Adverse Birth Outcomes Among Young Pregnant Women (<25 Years)**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("Images/AdjustedRatio.png")

@@

**Table 5:  Discriminative Performance (AUC) of Machine Learning Models Predicting Adverse Birth Outcomes Under Different IPTp Regimens**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::include_graphics("Images/Table6.Discriminative Performance (AUC) of ML Models.png")
